\vspace*{30mm}

# Supplementary Materials

# Supplementary Methods

## Moral Dilemma Wordings

:::{.NoIndent}
The wording for the "Baby" dilemma was as follows:
:::

> Enemy soldiers have taken over David's village. They have orders to kill all 
remaining civilians over the age of two. David and some of his townspeople have 
sought refuge in two rooms of the cellar of a large house. Outside, he hears the
voices of soldiers who have come to search the house for valuables. David's baby,
who is with him in the room, begins to cry loudly. David puts his hand over the
baby's mouth to block the sound. If David removes his hand from the baby's
mouth, its crying will summon the attention of the soldiers who will spare his 
baby's life, but will kill David and the others hiding in both rooms. To save 
himself and the others, David must keep his hand on the baby's mouth and smother
the baby to death.

:::{.NoIndent}
The wording for the "Bike" dilemma was as follows:
:::

> James is an expert motorcycle rider who has gone on vacation in order to 
participate in Bike Week. Thousands of other motorcycle riders from across the 
country have come to ride in this event. As he is riding down the road in front
of a large group of other riders, he sees that someone up ahead is losing
control of their bike. As he speeds up to pull alongside the unstable rider, he
realizes that this person is going to crash at any second. This would certainly
result in a large pile-up and several deaths as the riders behind James run over
each other trying to avoid the crashed rider. James realizes that he could
physically run this rider off the road and into some trees. This would cause the
other rider to crash and, at his current speed, almost certainly die, but it
would prevent a crash in the middle of the street and the large pile-up of 
riders.

## Specifications for Statistical Models

We used the *brms* R package to fit the Bayesian models in this paper. Below, we
outline the R code for the different models that we fitted.

Since the perceptions of the moral advisors (perceived trustworthiness, blame, 
trusting on other issues, surprise, and human-like perceptions) were measured on
1-7 Likert scales, we used the following cumulative-link ordinal model to 
analyse these outcome variables, including the interaction between advisor type 
and advice type:

```{r echo=TRUE, eval=FALSE}
brm(
  formula = response ~ 1 + advisor * advice + (1 | id) + 
    (1 + advisor * advice | country),
  data = data,
  family = cumulative,
  prior = c(
    prior(normal(0, 2), class = Intercept),
    prior(normal(0, 1), class = b),
    prior(exponential(2), class = sd),
    prior(lkj(2), class = cor)
  ),
  ...
)
```

When modelling participants' moral judgements on a 0-1 slider scale, we used a 
zero-one-inflated beta model, since this model is uniquely suited to the 
constraints of 0-1 bounded data (see 
[here](https://vuorre.com/posts/analyze-analog-scale-ratings-with-zero-one-inflated-beta-models/index.html)).
With this model, it is possible to simultaneously model the location and 
precision (phi) of the data within the bounds of 0 and 1, as well as the 
probability of zero-one-inflation (zoi) and the conditional probability of 
observing a zero or a one (coi). We take this approach here, allowing the time
period (pre vs. post) to interact with advisor type and advice type:

```{r echo=TRUE, eval=FALSE}
brm(
  formula = bf(
    judgement ~ 1 + time * advisor * advice + (1 | id) + 
      (1 + time * advisor * advice | country),
    phi ~ 1 + time * advisor * advice + (1 | id) + (1 | country),
    zoi ~ 1 + time * advisor * advice + (1 | id) + (1 | country),
    coi ~ 1 + time * advisor * advice + (1 | id) + (1 | country)
  ),
  data = data,
  family = zero_one_inflated_beta,
  prior = c(
    prior(normal(0, 1), class = Intercept),
    prior(normal(0, 1), class = Intercept, dpar = phi),
    prior(normal(0, 1), class = Intercept, dpar = zoi),
    prior(normal(0, 1), class = Intercept, dpar = coi),
    prior(normal(0, 1), class = b),
    prior(normal(0, 1), class = b, dpar = phi),
    prior(normal(0, 1), class = b, dpar = zoi),
    prior(normal(0, 1), class = b, dpar = coi),
    prior(exponential(2), class = sd),
    prior(exponential(2), class = sd, dpar = phi),
    prior(exponential(2), class = sd, dpar = zoi),
    prior(exponential(2), class = sd, dpar = coi),
    prior(lkj(3), class = cor)
  ),
  ...
)
```

Finally, when modelling participants' shift in confidence after seeing the
advice, we fitted the following cumulative-link ordinal model to the 1-7 Likert
scale data:

```{r echo=TRUE, eval=FALSE}
brm(
  formula = confidence ~ 1 + time * advisor * advice + (1 | id) + 
    (1 + time * advisor * advice | country),
  data = data,
  family = cumulative,
  prior = c(
    prior(normal(0, 2), class = Intercept),
    prior(normal(0, 1), class = b),
    prior(exponential(2), class = sd),
    prior(lkj(3), class = cor)
  ),
  ...
)
```

For more details about these models, and for variants of these models that 
include additional moderators, see here: 
<https://github.com/ScottClaessens/trustAMAs>

\newpage

# Supplementary Figures

::: {#suppfig-means-by-dilemma suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_by_dilemma)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across countries and split by dilemma. Box 
and whisker plots show the distributions of the data, with horizontal lines 
representing sample medians, boxes representing interquartile ranges, and 
whiskers representing ranges. Point ranges show model-estimated marginal means,
with points representing posterior medians and line ranges representing 95% 
highest posterior density intervals.

:::

\newpage

::: {#suppfig-means-by-order suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_by_order)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across dilemmas and countries and split by 
the first and second experimental blocks. Box and whisker plots show the 
distributions of the data, with horizontal lines representing sample medians, 
boxes representing interquartile ranges, and whiskers representing ranges. Point
ranges show model-estimated marginal means, with points representing posterior
medians and line ranges representing 95% highest posterior density intervals.

:::

\newpage

::: {#suppfig-means-control-judgement suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_control_judgement)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across dilemmas and countries and controlling
for participants' own pre-advice moral judgments. Box and whisker plots show 
the distributions of the data, with horizontal lines representing sample 
medians, boxes representing interquartile ranges, and whiskers representing 
ranges. Point ranges show model-estimated marginal means, with points 
representing posterior medians and line ranges representing 95% highest 
posterior density intervals.

:::

\newpage

::: {#suppfig-means-itt suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_itt)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across dilemmas and countries and including
participants who failed the comprehension checks ("intention to treat" analysis).
Box and whisker plots show the distributions of the data, with horizontal lines 
representing sample medians, boxes representing interquartile ranges, and 
whiskers representing ranges. Point ranges show model-estimated marginal means,
with points representing posterior medians and line ranges representing 95% 
highest posterior density intervals.

:::

\newpage

::: {#suppfig-ind-diffs suppfig-pos="H"}

```{r, fig.height=5, fig.width=5}
tar_read(plot7_combined)
```

Assessing variation in the experimental effects across individuals. The top row 
shows how the estimated difference in perceived trustworthiness (on a 1-7 Likert 
scale) between deontological and utilitarian advisors changes with participants' 
religiosity and political conservatism. The bottom row shows how the estimated 
difference in perceived trustworthiness (on a 1-7 Likert scale) between human 
and AI advisors changes with participants' familiarity with AI and frequency of 
AI use. Lines and shaded areas reflect posterior medians and 50% and 95% highest 
posterior density intervals. Dashed horizontal lines indicate no difference.

:::

\newpage

::: {#suppfig-country-slopes-blame suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_blame)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting blame, alongside overall 
fixed effects (red). Points and line ranges reflect posterior medians and 66% 
and 95% highest posterior density intervals, respectively. Densities reflect 
samples from the posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-trust-other-issues suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_trust_other_issues)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting trust on other issues, 
alongside overall fixed effects (red). Points and line ranges reflect posterior
medians and 66% and 95% highest posterior density intervals, respectively. 
Densities reflect samples from the posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-surprise suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_surprise)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting surprise, alongside overall
fixed effects (red). Points and line ranges reflect posterior medians and 66% 
and 95% highest posterior density intervals, respectively. Densities reflect 
samples from the posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-humanlike suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_humanlike)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting human-like perceptions, 
alongside overall fixed effects (red). Points and line ranges reflect posterior
medians and 66% and 95% highest posterior density intervals, respectively. 
Densities reflect samples from the posterior distribution.

:::

\newpage

::: {#suppfig-cross-cultural suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot8_combined)
```

Assessing variation in the experimental effects across different cultural 
variables. The top row shows how the estimated difference in perceived 
trustworthiness (on a 1-7 Likert scale) between deontological and utilitarian
advisors varies by relational mobility, tightness-looseness, and individualism.
The bottom row shows how the estimated difference in perceived trustworthiness
(on a 1-7 Likert scale) between human and AI advisors varies by government AI 
readiness and the global AI index. Lines and shaded areas reflect posterior 
medians and 50% and 95% credible intervals for visualization purposes. Letters 
are ISO-2 country codes. All models control for spatial and cultural 
non-independence between countries.

:::

\newpage

::: {#suppfig-judgment-shift-block1 suppfig-pos="H"}

```{r, fig.height=3, fig.width=6}
tar_read(plot10_judgement_shift)
```

Effects of moral advice on participant judgments in the first block only. 
Estimated pre-post shifts in judgments on a 0-1 sliding scale, for human and AI
advisors giving deontological and utilitarian advice. The grey dashed line 
indicates no pre-post shift in judgment. Points, line ranges, and densities 
reflect posterior medians, 95% highest posterior density intervals, and samples 
from the posterior distribution, respectively.

:::

\newpage

::: {#suppfig-judgment-shift-relational-mobility suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot11_relational_mobility_latent)
```

The relationship between relational mobility and pre-post shifts in moral 
judgments on a 0-1 sliding scale. Negative pre-post shifts indicate judgment 
updating towards the deontological option, positive pre-post shifts indicate 
judgment updating towards the utilitarian option. Lines and shaded areas reflect 
posterior medians and 50% and 95% credible intervals for visualization purposes. 
Points are country averages. Letters are ISO-2 country codes. The model controls
for spatial and cultural non-independence between countries.

:::

\newpage

::: {#suppfig-judgment-shift-tightness suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot11_tightness)
```

The relationship between cultural tightness and pre-post shifts in moral 
judgments on a 0-1 sliding scale. Negative pre-post shifts indicate judgment 
updating towards the deontological option, positive pre-post shifts indicate 
judgment updating towards the utilitarian option. Lines and shaded areas reflect 
posterior medians and 50% and 95% credible intervals for visualization purposes. 
Points are country averages. Letters are ISO-2 country codes. The model controls
for spatial and cultural non-independence between countries.

:::

\newpage

::: {#suppfig-judgment-shift-individualism suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot11_individualism)
```

The relationship between individualism and pre-post shifts in moral judgments on
a 0-1 sliding scale. Negative pre-post shifts indicate judgment updating towards
the deontological option, positive pre-post shifts indicate judgment updating 
towards the utilitarian option. Lines and shaded areas reflect posterior medians
and 50% and 95% credible intervals for visualization purposes. Points are 
country averages. Letters are ISO-2 country codes. The model controls for 
spatial and cultural non-independence between countries.

:::

\newpage

::: {#suppfig-judgment-shift-AI-readiness suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot11_AI_readiness)
```

The relationship between AI government readiness and pre-post shifts in moral 
judgments on a 0-1 sliding scale. Negative pre-post shifts indicate judgment 
updating towards the deontological option, positive pre-post shifts indicate 
judgment updating towards the utilitarian option. Lines and shaded areas reflect 
posterior medians and 50% and 95% credible intervals for visualization purposes. 
Points are country averages. Letters are ISO-2 country codes. The model controls
for spatial and cultural non-independence between countries.

:::

::: {#suppfig-judgment-shift-AI-index suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot11_AI_index)
```

The relationship between the global AI index and pre-post shifts in moral 
judgments on a 0-1 sliding scale. Negative pre-post shifts indicate judgment 
updating towards the deontological option, positive pre-post shifts indicate
judgment updating towards the utilitarian option. Lines and shaded areas reflect 
posterior medians and 50% and 95% credible intervals for visualization purposes. 
Points are country averages. Letters are ISO-2 country codes. The model controls
for spatial and cultural non-independence between countries.

:::

\newpage

::: {#suppfig-confidence suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot2_confidence)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice. Points and line ranges represent posterior medians and 50% and 95%
highest posterior density intervals, respectively.

:::

\newpage

::: {#suppfig-confidence-baby suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot4_confidence_baby)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice, for the Baby dilemma only. Points and line ranges represent 
posterior medians and 50% and 95% highest posterior density intervals, 
respectively.

:::

\newpage

::: {#suppfig-confidence-bike suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot4_confidence_bike)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice, for the Bike dilemma only. Points and line ranges represent 
posterior medians and 50% and 95% highest posterior density intervals, 
respectively.

:::

\newpage

::: {#suppfig-sample suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_sample)
```

Overall characteristics of the sample

:::

\newpage

::: {#suppfig-sample-representativeness suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_sample_representative)
```

Representativeness of the final sample. Bars show the observed proportions in 
each age-gender category, lines show the targeted proportions in our 
nationally representative quotas. Differences between observed and targeted 
proportions in each quota category were less than 10% for all countries except 
China, which was skewed slightly towards younger participants.

:::

\newpage

::: {#suppfig-prior-check suppfig-pos="H"}

```{r, fig.height=4, fig.width=4}
tar_read(plot1_prior_only)
```

Prior predictive check for the pre-registered models. These models analyse 
perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across dilemmas and countries. Point ranges 
show marginal means from the model before seeing the data, with points 
representing prior medians and line ranges representing 95% highest posterior 
density intervals.

:::

\newpage

::: {#suppfig-means-prior-sensitivity suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_prior_sensitivity)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian (blue) advice, pooling across dilemmas and countries and using a 
wider prior on the experimental effects - Normal(0, 2) instead of Normal(0, 1). 
Box and whisker plots show the distributions of the data, with horizontal lines 
representing sample medians, boxes representing interquartile ranges, and 
whiskers representing ranges. Point ranges show model-estimated marginal means, 
with points representing posterior medians and line ranges representing 95% 
highest posterior density intervals.

:::

\newpage

::: {#suppfig-trace-trustworthy suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot1_trace_trustworthy)
```

MCMC trace plot of post-warmup iterations for the model predicting 
trustworthiness ratings.

:::

\newpage

::: {#suppfig-trace-blame suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot1_trace_blame)
```

MCMC trace plot of post-warmup iterations for the model predicting blame 
ratings.

:::

\newpage

::: {#suppfig-trace-trust-other-issues suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot1_trace_trust_other_issues)
```

MCMC trace plot of post-warmup iterations for the model predicting "trust on 
other issues" ratings.

:::

\newpage

::: {#suppfig-trace-surprise suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot1_trace_surprise)
```

MCMC trace plot of post-warmup iterations for the model predicting surprise 
ratings.

:::

\newpage

::: {#suppfig-trace-humanlike suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot1_trace_humanlike)
```

MCMC trace plot of post-warmup iterations for the model predicting human-like 
vs. machine-like ratings.

:::

\newpage

::: {#suppfig-trace-judgements suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot2_trace)
```

MCMC trace plot of post-warmup iterations for the model predicting judgment 
updating. Note that only fixed effects from the beta component of the model are
presented here for space reasons.

:::

\newpage

::: {#suppfig-pp-check-ordinal suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot1_pp_check)
```

Posterior predictive checks for models predicting perceptions of the moral
advisors. Light blue bars represent observed data counts, dark blue point ranges
represent simulated data from the posterior predictive distribution, with points
reflecting posterior medians and line ranges reflecting 90% credible intervals.

:::

\newpage

::: {#suppfig-pp-check-judgements suppfig-pos="H"}

```{r, fig.height=4, fig.width=4.5}
tar_read(plot2_pp_check)
```

Posterior predictive check from the zero-one-inflated beta model predicting 
moral judgments on the 0-1 sliding scale. The dark blue density represents the 
distribution of the observed data. The light blue densities represent different 
draws from the posterior predictive distribution.

:::

\newpage

# Supplementary Tables

::: {.landscape}

::: {#supptbl-contrasts-by-dilemma supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_by_dilemma) %>%
  mutate(
    group = c(
      rep("Baby dilemma", 5),
      rep("Bike dilemma", 5)
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
countries and split by dilemma. Numbers reflect differences in marginal means on
the log-odds scale. The interaction effects represent the interaction between 
advisor type and advice type (i.e., the difference between the differences in 
the rows above). Main numbers are posterior medians, numbers in square brackets 
are 95% highest posterior density intervals.

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-by-order supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_by_order) %>%
  mutate(
    group = c(
      rep("Block 1", 5),
      rep("Block 2", 5)
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and split by the first and second experimental blocks.
Numbers reflect differences in marginal means on the log-odds scale. The 
interaction effects represent the interaction between advisor type and advice 
type (i.e., the difference between the differences in the rows above). Main 
numbers are posterior medians, numbers in square brackets are 95% highest 
posterior density intervals.

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-control-judgement supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_control_judgement) %>%
  mutate(
    group = c(
      rep("Effect of advice type", 2),
      rep("Effect of advisor type", 2),
      "Interaction effect"
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and controlling for participants' pre-advice moral 
judgments. Numbers reflect differences in marginal means on the log-odds scale.
The interaction effects represent the interaction between advisor type and 
advice type (i.e., the difference between the differences in the rows above). 
Main numbers are posterior medians, numbers in square brackets are 95% highest 
posterior density intervals.

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-itt supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_itt) %>%
  mutate(
    group = c(
      rep("Effect of advice type", 2),
      rep("Effect of advisor type", 2),
      "Interaction effect"
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and including participants who failed the comprehension 
checks ("intention to treat" analysis). Numbers reflect differences in marginal
means on the log-odds scale. The bottom row represents the interaction between
advisor type and advice type (i.e., the difference between the differences in
the rows above). Main numbers are posterior medians, numbers in square brackets
are 95% highest posterior density intervals.

:::

:::

::: {.landscape}

::: {#supptbl-model11 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table11) %>%
  flextable() %>%
  theme_apa() %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  fontsize(size = 10) %>%
  line_spacing() %>%
  autofit()
```

Interaction effects for different country-level variables predicting pre-post
shifts in judgment updating on a 0-1 sliding scale. Numbers reflect interaction
effects on the log-odds scale. Main numbers are posterior medians, numbers in 
square brackets are 95% highest posterior density intervals.

:::

:::

::: {.landscape}

::: {#supptbl-sample-by-country supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_sample) %>%
  flextable() %>%
  theme_apa() %>%
  fontsize(size = 9) %>%
  line_spacing() %>%
  autofit()
```

Breakdown of the sample by country

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-prior-sensitivity supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_prior_sensitivity) %>%
  mutate(
    group = c(
      rep("Effect of advice type", 2),
      rep("Effect of advisor type", 2),
      "Interaction effect"
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and using a wider prior on the experimental effects - 
Normal(0, 2) instead of Normal(0, 1). Numbers reflect differences in marginal
means on the log-odds scale. The bottom row represents the interaction between
advisor type and advice type (i.e., the difference between the differences in
the rows above). Main numbers are posterior medians, numbers in square brackets
are 95% highest posterior density intervals.

:::

:::
