\vspace*{30mm}

# Supplementary Materials

# Moral Dilemma Wordings

:::{.NoIndent}
The wording for the "Baby" dilemma was as follows:
:::

> Enemy soldiers have taken over David's village. They have orders to kill all 
remaining civilians over the age of two. David and some of his townspeople have 
sought refuge in two rooms of the cellar of a large house. Outside, he hears the
voices of soldiers who have come to search the house for valuables. David's baby,
who is with him in the room, begins to cry loudly. David puts his hand over the
baby's mouth to block the sound. If David removes his hand from the baby's
mouth, its crying will summon the attention of the soldiers who will spare his 
baby's life, but will kill David and the others hiding in both rooms. To save 
himself and the others, David must keep his hand on the baby's mouth and smother
the baby to death.

:::{.NoIndent}
The wording for the "Bike" dilemma was as follows:
:::

> James is an expert motorcycle rider who has gone on vacation in order to 
participate in Bike Week. Thousands of other motorcycle riders from across the 
country have come to ride in this event. As he is riding down the road in front
of a large group of other riders, he sees that someone up ahead is losing
control of their bike. As he speeds up to pull alongside the unstable rider, he
realizes that this person is going to crash at any second. This would certainly
result in a large pile-up and several deaths as the riders behind James run over
each other trying to avoid the crashed rider. James realizes that he could
physically run this rider off the road and into some trees. This would cause the
other rider to crash and, at his current speed, almost certainly die, but it
would prevent a crash in the middle of the street and the large pile-up of 
riders.

\newpage

# Supplementary Figures

::: {#suppfig-means-by-dilemma suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_by_dilemma)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian advice (blue), pooling across countries and split by dilemma. Box 
and whisker plots show the distributions of the data, with horizontal lines 
representing sample medians, boxes representing interquartile ranges, and 
whiskers representing ranges. Point ranges show model-estimated marginal means,
with points representing posterior medians and line ranges representing 95% 
credible intervals.

:::

\newpage

::: {#suppfig-means-by-order suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_by_order)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian advice (blue), pooling across dilemmas and countries and split by 
the first and second experimental blocks. Box and whisker plots show the 
distributions of the data, with horizontal lines representing sample medians, 
boxes representing interquartile ranges, and whiskers representing ranges. Point
ranges show model-estimated marginal means, with points representing posterior
medians and line ranges representing 95% credible intervals.

:::

\newpage

::: {#suppfig-means-itt suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot_means_itt)
```

Perceptions of AI and human moral advisors giving deontological (orange) and 
utilitarian advice (blue), pooling across dilemmas and countries and including
participants who failed the comprehension checks ("intention to treat" analysis).
Box and whisker plots show the distributions of the data, with horizontal lines 
representing sample medians, boxes representing interquartile ranges, and 
whiskers representing ranges. Point ranges show model-estimated marginal means,
with points representing posterior medians and line ranges representing 95% 
credible intervals.

:::

\newpage

::: {#suppfig-ind-diffs suppfig-pos="H"}

```{r, fig.height=5, fig.width=5}
tar_read(plot7_combined)
```

Variation in the experimental effects across individuals. The top row shows how
the estimated difference in trust (on a 1-7 Likert scale) between deontological
and utilitarian advisors changes with participants' religiosity and political 
conservatism. The bottom row shows how the estimated difference in trust (on a 
1-7 Likert scale) between human and AI advisors changes with participants' 
familiarity with AI and frequency of AI use. Lines and shaded areas reflect 
posterior medians and 50% and 95% credible intervals. Dashed horizontal lines 
indicate no difference.

:::

\newpage

::: {#suppfig-country-slopes-blame suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_blame)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting blame, alongside overall 
fixed effects (red). Points and line ranges reflect posterior medians and 66% 
and 95% credible intervals, respectively. Densities reflect samples from the 
posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-trust-other-issues suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_trust_other_issues)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting trust on other issues, 
alongside overall fixed effects (red). Points and line ranges reflect posterior
medians and 66% and 95% credible intervals, respectively. Densities reflect 
samples from the posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-surprise suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_surprise)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting surprise, alongside overall
fixed effects (red). Points and line ranges reflect posterior medians and 66% 
and 95% credible intervals, respectively. Densities reflect samples from the 
posterior distribution.

:::

\newpage

::: {#suppfig-country-slopes-humanlike suppfig-pos="H"}

```{r, fig.height=4, fig.width=9}
tar_read(plot1_humanlike)
```

Variation in model parameters across countries. Estimates are the varying 
log-odds slopes from our multilevel model predicting human-like perceptions, 
alongside overall fixed effects (red). Points and line ranges reflect posterior
medians and 66% and 95% credible intervals, respectively. Densities reflect 
samples from the posterior distribution.

:::

\newpage

::: {#suppfig-cross-cultural suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot8_combined)
```

Variation in the experimental effects across different cultural variables. The 
top row shows how the estimated difference in trust (on a 1-7 Likert scale) 
between deontological and utilitarian advisors varies by relational mobility, 
tightness-looseness, and individualism. The bottom row shows how the estimated 
difference in trust (on a 1-7 Likert scale) between human and AI advisors varies
by government AI readiness and the global AI index. Lines and shaded areas 
reflect posterior medians and 50% and 95% credible intervals. Letters are ISO-2
country codes. All models control for spatial and cultural non-independence 
between countries.

:::

\newpage

::: {#suppfig-judgment-shift-block1 suppfig-pos="H"}

```{r, fig.height=3, fig.width=6}
tar_read(plot10_judgement_shift)
```

Effects of moral advice on participant judgments in the first block only. 
Estimated pre-post shifts in judgments on a 0-1 sliding scale, for human and AI
advisors giving deontological and utilitarian advice. The grey dashed line 
indicates no pre-post shift in judgment. Points, line ranges, and densities 
reflect posterior medians, 95% credible intervals, and samples from the 
posterior distribution, respectively.

:::

\newpage

::: {#suppfig-confidence suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot2_confidence)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice. Points and line ranges represent posterior medians and 50% and 95%
credible intervals, respectively.

:::

\newpage

::: {#suppfig-confidence-baby suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot4_confidence_baby)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice, for the Baby dilemma only. Points and line ranges represent 
posterior medians and 50% and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-confidence-bike suppfig-pos="H"}

```{r, fig.height=4.5, fig.width=6}
tar_read(plot4_confidence_bike)
```

Pre-post increase in participants' confidence in their judgment after seeing the
moral advice, for the Bike dilemma only. Points and line ranges represent 
posterior medians and 50% and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-sample suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_sample)
```

Overall characteristics of the sample

:::

\newpage

::: {#suppfig-sample-representativeness suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_sample_representative)
```

Representativeness of the final sample. Bars show the observed proportions in 
each age-gender category, lines show the targeted proportions in our 
nationally representative quotas. Differences between observed and targeted 
proportions in each quota category were less than 10% for all countries except 
China, which was skewed slightly towards younger participants.

:::

\newpage

# Supplementary Tables

::: {.landscape}

::: {#supptbl-contrasts-by-dilemma supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_by_dilemma) %>%
  mutate(
    group = c(
      rep("Baby dilemma", 5),
      rep("Bike dilemma", 5)
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
countries and split by dilemma. Numbers reflect differences in marginal means on
the log-odds scale. The interaction effects represent the interaction between 
advisor type and advice type (i.e., the difference between the differences in 
the rows above). Main numbers are posterior medians, numbers in square brackets 
are 95% credible intervals.

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-by-order supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_by_order) %>%
  mutate(
    group = c(
      rep("Block 1", 5),
      rep("Block 2", 5)
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and split by the first and second experimental blocks.
Numbers reflect differences in marginal means on the log-odds scale. The 
interaction effects represent the interaction between advisor type and advice 
type (i.e., the difference between the differences in the rows above). Main 
numbers are posterior medians, numbers in square brackets are 95% credible 
intervals.

:::

:::

::: {.landscape}

::: {#supptbl-contrasts-itt supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_pairwise_contrasts_itt) %>%
  mutate(
    group = c(
      rep("Effect of advice type", 2),
      rep("Effect of advisor type", 2),
      "Interaction effect"
      )
    ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 9)
```

Pairwise contrasts for perceptions of the moral advisors, pooling across 
dilemmas and countries and including participants who failed the comprehension 
checks ("intention to treat" analysis). Numbers reflect differences in marginal
means on the log-odds scale. The bottom row represents the interaction between
advisor type and advice type (i.e., the difference between the differences in
the rows above). Main numbers are posterior medians, numbers in square brackets
are 95% credible intervals.

:::

:::

::: {.landscape}

::: {#supptbl-sample-by-country supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_sample) %>%
  flextable() %>%
  theme_apa() %>%
  fontsize(size = 9) %>%
  line_spacing() %>%
  autofit()
```

Breakdown of the sample by country

:::

:::
