---
title: "Trust in artificial moral advisors across cultures"
shorttitle: "Trust in artificial moral advisors"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Konrad Bocian
    corresponding: false
    orcid: 0000-0002-8652-0167
    affiliations:
      - name: SWPS University
        department: Department of Psychology in Sopot
    role:
      - methodology
      - editing
  - name: Paulo Sérgio Boggio
    corresponding: false
    orcid: 0000-0002-6109-0447
    affiliations:
      - name: Mackenzie Presbyterian University
        department: Social and Cognitive Neuroscience Laboratory
      - name: National Institute of Science and Technology on Social and Affective Neuroscience (INCT-SANI)
    role:
      - methodology
      - editing
  - name: Grégory Fiorio
    corresponding: false
    orcid: 0009-0008-3731-5678
    affiliations:
      - name: Université PSL
        department: Institut Jean Nicod, Département d’Études Cognitives, École Normale Supérieure
    role:
      - methodology
      - editing
  - name: Léo Fitouchi
    corresponding: false
    orcid: 0000-0002-1724-0129
    affiliations:
      - name: Toulouse School of Economics
        department: Institute for Advanced Study in Toulouse, Department of Social and Behavioral Sciences
    role:
      - methodology
      - editing
  - name: Zeynep Genç
    corresponding: false
    orcid: 0009-0004-8586-5771
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - methodology
      - editing
  - name: Ivar R. Hannikainen
    corresponding: false
    orcid: 0000-0003-0623-357X
    affiliations:
      - name: University of Granada
        department: Department of Philosophy
    role:
      - methodology
      - editing
  - name: Lea C. Kamitz
    corresponding: false
    orcid: 0000-0003-1290-5088
    affiliations:
      - name: Anglia Ruskin University
        department: IPPPRI
    role:
      - methodology
      - editing
  - name: Tamino Konur
    corresponding: false
    orcid: 0009-0005-4706-1200
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - methodology
      - editing
  - name: Ethan Landes
    corresponding: false
    orcid: 0000-0002-1186-1717
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - methodology
      - editing
  - name: Peng Liu
    corresponding: false
    orcid: 0000-0003-4929-0531
    affiliations:
      - name: Zhejiang University
        department: Center for Psychological Sciences
      - name: National University of Singapore
        department: Centre for Biomedical Ethics, Yong Loo Lin School of Medicine
    role:
      - methodology
      - editing
  - name: Katarzyna Miazek
    corresponding: false
    orcid: 0000-0001-9072-2688
    affiliations:
      - name: SWPS University
        department: Department of Psychology in Sopot
    role:
      - methodology
      - editing
  - name: Waldir M. Sampaio
    corresponding: false
    orcid: 0000-0002-6066-4314
    affiliations:
      - name: Pontifical Catholic University of Rio De Janeiro
        department: Department of Psychology
      - name: National Institute of Science and Technology on Social and Affective Neuroscience (INCT-SANI)
    role:
      - methodology
      - editing
  - name: Jorge Suárez
    corresponding: false
    orcid: 0000-0002-6215-2064
    affiliations:
      - name: University of Granada
        department: Department of Philosophy
    role:
      - methodology
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - methodology
      - editing
  - name: Onurcan Yilmaz
    corresponding: false
    orcid: 0000-0002-6094-7162
    affiliations:
      - name: Kadir Has University
        department: Department of Psychology
    role:
      - methodology
      - editing
  - name: Hongbo Yu
    corresponding: false
    orcid: 0000-0002-3384-7772
    affiliations:
      - name: University of California Santa Barbara
        department: Department of Psychological and Brain Sciences
    role:
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: United Kingdom
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - writing
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed and may differ from the final version.
    data-sharing: Word count (excl. abstract, methods, figure legends, and references) = 5,112 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "Developers of artificial intelligence are already building prototypes for artificial moral advisors: autonomous systems designed to provide humans with recommendations on ethical issues. Yet it remains unclear whether people will actually trust and adopt these technologies. Here, we investigate perceptions of human and artificial moral advisors in a large-scale cross-cultural experiment (12 countries; *N* = 6,896). We show that people trust artificial advisors less than human ones, especially when their advice aligns with utilitarian rather than deontological principles. Nonetheless, people update their own moral judgments in the direction of the advice -- regardless of whether it came from a human or AI advisor -- and do so more when the advice is deontological. These findings reveal a psychological paradox: people trust artificial moral advisors less, yet still defer to them in practice. This dissociation raises serious questions about the potential influence of machines on human morality, with implications for AI governance."
keywords: [artificial intelligence, morality, trust, moral influence]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-docx:
    fig-dpi: 300
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
csl: nature.csl
---

```{r setup}
library(flextable)
library(patchwork)
library(rethinking)
library(targets)
library(tidyverse)

# rounding function to use throughout (2 decimal places by default)
roundd <- function(x, digits = 2) {
  as.character(format(round(x, digits), nsmall = digits))
}
```

Artificial intelligence (AI) – any system that uses an algorithm or a 
statistical model to perform tasks that usually require human intelligence – is 
reshaping the world around us. By leveraging the computational power of AI 
agents, we can build on our own skills and therefore perform tasks more easily, 
more quickly, and perhaps even better than we could alone. Students use 
large language models (LLMs) to help them write assignments, banks rely on AI to
help with credit checks, and hospitals use AI tools to provide diagnoses for 
diseases. But could -- and should -- AI also be used to help us make better
moral decisions [@Giubilini2018; @Lara2020; @Sinnott2021]?

Artificial moral advisors (AMAs) refer to AI systems that could assist humans 
in making ethical decisions, leveraging artificial intelligence to analyse moral
dilemmas, engage in Socratic dialogue about moral matters [@Lara2020], and/or 
provide recommendations based on established ethical theories, principles, and 
guidelines [@Giubilini2018]. The idea, at root, is simple: AI provides a tool 
that can help enhance human activity in many different domains, so why not use 
AI to help people make better moral decisions? Exactly what "better" means in 
the moral domain is up for debate [@Vallor2024]. But the general argument is 
that AMAs could serve a function akin to the "ideal observer" [@Firth1952] by 
offering dispassionate and consistent judgments free from human biases 
[@Giubilini2018; @Sinnott2021]. Over the last decade, there has been growing 
philosophical attention to the possible future role of AMAs [@Giubilini2018; 
@Landes2025; @Lara2020; @Liu2022; @Sinnott2021], and now with the release of 
openly available LLMs like ChatGPT, Gemini, and Claude, this interest has 
rapidly moved from hypothetical to practical.

While contemporary LLMs like ChatGPT are not designed specifically to assist in 
morality, their flexibility and accessibility means that they do already give 
advice on moral dilemmas [@Krugel2023]. Perhaps concerningly, ethical advice 
surreptitiously written by ChatGPT is rated as more moral, thoughtful, and 
correct than advice written by a human moral expert [@Dillion2025]. Beyond 
general-purpose LLMs, companies are already working on prototypes for AI-powered
systems designed specifically to model morality. For example, the Allen 
Institute's Delphi [@Jiang2025] is trained on a large corpus of moral judgments,
and it is now even possible to converse with an AI chatbot trained on the 
writings of the renowned ethicist Peter Singer (<https://www.petersinger.ai/>).

Questions about how we should create AMAs, whether we should at all, and the 
long-term consequences of these technologies remain very much open [@Lara2020; 
@Landes_preprint; @Liu2022]. However, even if we set aside the crucial question
of whether people *should* trust AMAs, it remains unclear whether people will 
actually trust and adopt AMAs in practice. We can broadly define trust as 
"accept[ing] vulnerability based upon positive expectations of the intentions or
behavior of another" [@Rousseau1998], where this trust is typically understood 
as being driven by a perception of the agent's trustworthiness. While some 
philosophers have argued that concepts like trust and trustworthiness should 
not be applied to AI since machines lack the necessary agential capacities to be
genuinely trusted [@Metzinger2019], others have argued that people can and do 
trust AI in a manner akin to trust in other humans [@Lalot2024; @Malle2021]. But
how this trust will apply to people's interactions with AMAs is far from clear.
Even if people report that they trust AMAs, there is no guarantee that they
would listen to the moral advice by changing their judgment. Indeed, theoretical
models of trust in AI often distinguish between attitudinal trust and its 
antecedents, such as the perceived trustworthiness of the AI system, and 
behavioral consequences, such as willingness to use or rely on the technology in
practice [@Hoff2015; @Lalot2024; @Lee2004; @Schlicker2025; @Vereschak2021; 
@Wischnewski2023]. Given the increasing role that AI plays in life, and in 
particular claims that AI-powered moral advisors could help us make better moral
judgments, we need to understand whether people trust AMAs, whether this trust
depends on the kinds of moral decisions that the AMAs make, and whether people
will actually defer to AMAs on moral issues. Our paper addresses all these 
questions.

On the question of whether people will trust AMAs, a large body of research, 
mostly conducted prior to the release of contemporary LLMs, has documented the 
phenomenon of algorithm aversion: the tendency for individuals to distrust AI 
relative to humans, even when the AI boasts identical or even superior 
performance [@Dawes1979; @Dietvorst2015; @Meehl1954; @Meehl1957]. This algorithm
aversion leads to a specific distrust in AI making moral decisions, driven by
the perception that AI lacks internal experience [@Bigman2018]. These results 
suggest that people will distrust AI models that give advice on moral issues. 
However, these findings may no longer generalize in a post-ChatGPT era in which
people are interacting with LLMs on a regular basis. Moreover, the limited 
research there has been on trust in AI in the moral domain has tended to focus 
narrowly on Western populations [@Henrich2010], while moral norms, prevalence of
AI, and morally relevant cultural values vary dramatically across societies 
[@Barnes2024; @Dang2025; @Inglehart2000; @Jackson2024]. It therefore remains 
unclear whether people around the world will trust AMAs and whether levels of 
trust will vary across cultures.

There is reason to believe that people will trust AMAs differently depending on 
the moral decisions that they make. For AMAs to be trustworthy, they must be 
aligned with some normative standard, and there is unsurprisingly a great deal 
of interest in how AI can and should be aligned with our moral values -- and 
which ones specifically. A fundamental challenge that developers of AMAs will 
face is determining which kind of ethical framework to benchmark against, 
especially in moral dilemmas where different ethical frameworks endorse mutually
exclusive actions. For example, is it morally acceptable to break normal 
prohibitions against murder to prevent harm to a greater number? 
Consequentialist theories such as utilitarianism focus on the 'greatest good for 
the greatest number', positing that only consequences matter when making moral 
decisions [@Bentham1983; @Mill1863; @Singer1993]. In contrast, non-utilitarian 
deontological theories claim that we also have to consider rights, duties, and 
obligations when making moral judgments: for example, even if murder might bring
about good consequences, it should still be judged as wrong since we have a 
moral duty not to harm others [@Fried1978; @Kant2002; @Ross1930]. AMAs could 
theoretically provide advice in line with either of these philosophical 
frameworks, but the answers they give in dilemmas at the fault lines of
our morality are likely to shape how much they are trusted. A growing body of 
work suggests that people who give characteristically deontological responses to
sacrificial moral dilemmas (by rejecting instrumental harm) are seen as more 
trustworthy than those who give characteristically utilitarian responses (by 
endorsing instrumental harm) [@Brown2019; @Crockett2021; @Everett2016; 
@Everett2018; @Everett2021; @Rom2017; @Sacco2017], in part because people who 
make deontological decisions are perceived as warmer, more predictable, and more 
committed to their social partners [@Everett2016; @Everett2018; @Kreps2014; 
@Rom2017; @Turpin2021; @Uhlmann2013]. Yet preferences for deontological 
decision-makers can also be sensitive to the type of role one is judging, with
more acceptability for utilitarian decision-making from political leaders in
impartial roles [@Everett2018]. Given that AMAs are proposed to be able to 
provide more impartial advice in line with the idea of an ideal observer, it is 
therefore possible that people might prefer more utilitarian advice from AMAs.

Of most relevance to this paper, Myers and Everett [@Myers2025] recently 
conducted four pre-registered studies looking at how people trust AMAs. The
results showed that participants trusted AMAs less than human advisors and 
trusted AMAs less if they endorsed instrumental harm in line with utilitarian
philosophy. While these results are important in helping us understand how AMAs 
are differently trusted based on the kinds of decisions they make, there is 
still much we do not know. First, Myers and Everett [@Myers2025] only looked at 
participants in the UK, limiting cross-cultural generalizability. While 
cross-cultural research has shown that countries differ in their cooperation 
with AI [@Karpus2025], fears of it occupying various positions [@Dong2024], and 
legal concerns [@Ikkatai2024], it remains unclear how people across different 
countries that vary across different cultural indices perceive different kinds 
of AMAs. Second, Myers and Everett [@Myers2025] examined ratings of 
trustworthiness, willingness to trust the AI on other issues, and expected 
agreement in the future; however, we do not know if people will actually follow 
the advice from AMAs.

Even if people report that they trust AMAs, this stated trust may not lead 
people to defer to the moral advice from AMAs in practice: perceived 
trustworthiness does not always entail trusting behavior [@Hoff2015; 
@Schlicker2025; @Vereschak2021; @Wischnewski2023]. However, there is increasing 
evidence that people can be persuaded by contemporary consumer-grade LLMs in 
small but significant ways [@Schoenegger2025]. When directed to argue for a 
point in the non-moral domain, LLMs can change political beliefs [@Durmus2024; 
@Hackenburg2024; @Hackenburg2025], such as views on smoking bans and assault 
weapons bans [@Bai2025], and reduce endorsement of conspiracy theories 
[@Costello2024]. There is less work on AI persuasion in the moral domain, but 
some research suggests that American participants listen to pre-generated LLM 
advice in sacrificial moral dilemmas by giving judgments in line with the advice
provided [@Krugel2023] and other work suggests that British participants update
their moral judgments in everyday moral dilemmas when given advice from LLMs 
[@Landes_under_review]. This raises the important question of whether people 
will listen to AMAs less than humans and whether this will vary depending on the
type of advice (deontological or utilitarian) given by the AMA system.

In this research, we explore whether people trust AMAs compared to human moral 
advisors, whether this trust differs based on the kind of moral advice that the 
AMA gives in a sacrificial moral dilemma, and whether people update their own 
moral judgments when provided with advice from AMAs. Using a large 
cross-cultural sample (@fig-study-overview), we explore the generalizability of
these effects across individuals and populations.

```{r, fig.height=4, fig.width=7}
#| label: fig-study-overview
#| fig-cap: Overview of the study
#| fig-align: center
#| apa-note: (a) The flow of the experiment. We presented participants with a sacrificial moral dilemma (either the Bike dilemma or the Baby dilemma; within-subjects) and asked participants for their own moral judgment and their confidence in their judgment. We then presented participants with deontological or utilitarian advice (within-subjects) from either a human or an AI advisor (between-subjects). The advice was followed by several questions about the perceptions of the advisor, such as trust, and questions about participants’ revised moral judgment and confidence. Participants then repeated this flow for the other moral dilemma and the other direction of advice. (b) Examples of the deontological (orange) and utilitarian (blue) advice given by human and AI advisors in the study. (c) Countries included in the sample.

tar_read(plot_overview)
```

<br>

The flow of the experiment is shown in @fig-study-overview. We presented 
participants with one of two sacrificial dilemmas reflecting a tension between 
utilitarian and deontological theories about the acceptability of harm to one to
save a greater number of others (the “Bike” dilemma and the “Baby” dilemma; see
Methods for more details). After asking participants what they thought should be
done in the moral dilemma, we then presented participants with moral advice from
an advisor that was described as either a human or an AI moral expert 
(between-subjects). The advisor either gave characteristically utilitarian 
advice to endorse the harm, justifying this in terms of maximizing the greater 
good, or gave characteristically deontological advice to reject the harm, 
appealing to certain moral actions being wrong even if they have good 
consequences. We asked participants to rate various perceptions of the advisor, 
such as trustworthiness. Participants were then presented with the dilemma 
again, and asked if, having seen this advice, they wanted to update their 
initial judgment. Finally, participants repeated this flow for the other 
sacrificial dilemma and the other direction of advice: for example, if 
participants initially saw an advisor give deontological advice for the “Bike" 
dilemma, they then saw a different advisor give utilitarian advice for the 
“Baby” dilemma. We conducted the experiment in 12 countries (Brazil, Chile, 
China, France, Germany, India, Mexico, Poland, South Africa, Turkey, UK, USA; 
overall *N* = 6,896) chosen to maximize variation in geographic spread, cultural
backgrounds, and levels of AI progress according to global indices. In our 
pre-registration, we hypothesized that participants would trust AMAs less than 
human advisors, and utilitarian advisors less than deontological advisors 
(<https://osf.io/rcufb>).

# Results

## Perceptions of the Moral Advisors

We first looked at participants' overall perceptions of the moral advisors, 
pooling across dilemmas and countries (@fig-means; @tbl-contrasts). We found 
that the type of advice given (deontological vs. utilitarian) influenced 
participants' perceptions of the advisors. In line with our pre-registered 
predictions, advisors who gave utilitarian advice in the sacrificial dilemmas 
were perceived as less trustworthy than advisors who gave deontological advice.
Participants reported that they would blame someone else more for following 
utilitarian advice. In contrast to previous work [@Myers2025], we did not find 
that people were more surprised when the AI gave deontological advice rather 
than utilitarian advice. However, we found that advisors were perceived as more 
machine-like when they gave utilitarian advice, suggesting at least some belief 
about the association of machines with utilitarian decision making.

```{r, fig.height=5, fig.width=7}
#| label: fig-means
#| fig-cap: Perceptions of AI and human moral advisors giving deontological (orange) and utilitarian (blue) advice, pooling across dilemmas and countries
#| fig-align: center
#| apa-note: Box and whisker plots show the distributions of the data, with horizontal lines representing sample medians, boxes representing interquartile ranges, and whiskers representing ranges. Point ranges show model-estimated marginal means, with points representing posterior medians and line ranges representing 95% highest posterior density intervals.

tar_read(plot_means)
```

::: {.landscape}

```{r}
#| label: tbl-contrasts
#| tbl-cap: Pairwise contrasts for perceptions of the moral advisors, pooling across dilemmas and countries
#| apa-note: Numbers reflect differences in marginal means on the log-odds scale. For reference, on the log-odds scale, 0.23 is considered a small effect size, 0.54 a medium effect size, and 0.83 a large effect size [@Chen2010]. The bottom row represents the interaction between advisor type and advice type (i.e., the difference between the differences in the rows above). Main numbers are posterior medians, numbers in square brackets are 95% highest posterior density intervals.

tar_read(table_pairwise_contrasts) %>%
  mutate(
    group = c(
        rep("Effect of advice type", 2),
        rep("Effect of advisor type", 2),
        "Interaction effect"
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.7, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 10)
```

:::

In addition, we found that the type of advisor (human vs. AI) influenced 
participants' perceptions. In line with our pre-registered predictions and with 
previous work on algorithm aversion in the moral domain [@Bigman2018], 
artificial moral advisors were perceived as less trustworthy than human 
advisors. Participants reported that they would blame someone else more for 
following moral advice from an AI. Human advisors were seen as more human-like 
than AI advisors, especially when they gave deontological advice. Participants 
were also more surprised when deontological advice came from an AI, rather than
from a human.

In addition to our main pre-registered analyses, we assessed the robustness of 
these effects in several ways. First, we looked at the results split across the
two sacrificial dilemmas (@suppfig-means-by-dilemma; 
@supptbl-contrasts-by-dilemma). We found the same broad pattern of results for 
both dilemmas, though the type of advice (deontological vs. utilitarian) had a 
stronger influence on perceptions in the Baby dilemma compared to the Bike 
dilemma. Second, we looked at the results split by the order of the two blocks, 
finding little evidence of order effects (@suppfig-means-by-order; 
@supptbl-contrasts-by-order). Third, we included participants' pre-advice moral 
judgments as a covariate in the models to ensure that people's own moral 
convictions prior to receiving the advice were not confounding the results. The 
results were unchanged when controlling for participants' own moral judgments 
(@suppfig-means-control-judgement; @supptbl-contrasts-control-judgement). 
Fourth, we determined whether the effects held in an "intention-to-treat" 
analysis, finding that our results did not differ when we included participants
that we had excluded due to comprehension failures (@suppfig-means-itt; 
@supptbl-contrasts-itt).

## Variation across Individuals and Countries

```{r load-int-effects1}
tar_load(
  c(
    interaction_effect_religiosity,
    interaction_effect_political_ideology,
    interaction_effect_AI_familiarity,
    interaction_effect_AI_frequency,
    interaction_effect_tightness
  )
)

# reverse direction when predicting preference for deontological advisor
interaction_effect_religiosity        <- -interaction_effect_religiosity
interaction_effect_political_ideology <- -interaction_effect_political_ideology
interaction_effect_tightness          <- -interaction_effect_tightness
```

While overall effects capture broad trends, it is possible that our experimental 
manipulations worked differently for individuals with different demographics and 
for countries with different cultural backgrounds. To investigate this potential
heterogeneity, we ran several exploratory models including individual-level and
cultural-level variables as moderators of the experimental effects, focusing on 
the perceived trustworthiness of the moral advisors as the outcome.

We first explored whether the experimental effects varied by demographics and 
other individual-level characteristics. We allowed participants' religiosity and 
political ideology to moderate the effect of advice type (deontological vs. 
utilitarian) and allowed participants' familiarity with AI and frequency of AI 
use to moderate the effect of advisor type (human vs. AI). However, the 
experimental effects did not vary across these individual-level variables 
(@suppfig-ind-diffs). In particular, the effect of advice type on 
trustworthiness did not vary by religiosity (interaction effect = 
`r roundd(median(interaction_effect_religiosity))`, 95% highest posterior 
density interval 
[`r roundd(HPDI(interaction_effect_religiosity, prob = 0.95)[[1]])` 
`r roundd(HPDI(interaction_effect_religiosity, prob = 0.95)[[2]])`])
or political ideology (interaction effect = 
`r roundd(median(interaction_effect_political_ideology))`, 95% HPDI 
[`r roundd(HPDI(interaction_effect_political_ideology, prob = 0.95)[[1]])` 
`r roundd(HPDI(interaction_effect_political_ideology, prob = 0.95)[[2]])`])
and the effect of advisor type did not vary by familiarity with AI (interaction
effect = 
`r roundd(median(interaction_effect_AI_familiarity))`, 95% HPDI 
[`r roundd(HPDI(interaction_effect_AI_familiarity, prob = 0.95)[[1]])` 
`r roundd(HPDI(interaction_effect_AI_familiarity, prob = 0.95)[[2]])`]) 
or frequency of AI usage (interaction effect = 
`r roundd(median(interaction_effect_AI_frequency))`, 95% HPDI 
[`r roundd(HPDI(interaction_effect_AI_frequency, prob = 0.95)[[1]])` 
`r roundd(HPDI(interaction_effect_AI_frequency, prob = 0.95)[[2]])`]).

We also explored whether the experimental effects varied across countries with 
different cultural backgrounds. Extracting the varying slopes from our 
multilevel model, we found that the experimental effects of advisor type 
(human vs. AI) and advice type (deontological vs. utilitarian) on perceived 
trustworthiness were relatively robust across countries 
(@fig-country-slopes-trust; see Supplementary Figures 
[-@suppfig-country-slopes-blame] - [-@suppfig-country-slopes-humanlike] for 
other perceptions variables). However, there were some notable exceptions -- the
95% HPDIs for the effect of advisor type included zero for India, Chile, China, 
and Mexico, and the 95% HPDIs for the effect of advice type included zero for 
India, China, and the USA. To explain the variation across countries, we 
explored whether the experimental effects varied by cultural factors highlighted
by previous research [@Awad2020; @Graham2016]. Accounting for the geographic and 
cultural non-independence of countries [@Claessens2023], we included global 
indices of AI progress and government readiness as moderators of the effect of 
advisor type (human vs. AI) and relational mobility, cultural 
tightness-looseness, and individualism as moderators of the effect of advice 
type (deontological vs. utilitarian). Of these cultural factors, we only found 
evidence that the experimental effects varied by tightness-looseness 
(@fig-tightness). Looser countries, like Brazil, Mexico, and Chile, trusted the 
deontological advisor relatively more than tighter countries like India and 
China (interaction effect = 
`r roundd(median(interaction_effect_tightness))`, 95% HPDI 
[`r roundd(HPDI(interaction_effect_tightness, prob = 0.95)[[1]])` 
`r roundd(HPDI(interaction_effect_tightness, prob = 0.95)[[2]])`]). Other 
cultural factors did not moderate the experimental effects 
(@suppfig-cross-cultural).

```{r, fig.height=4, fig.width=6.5}
#| label: fig-country-slopes-trust
#| fig-cap: Variation in model parameters across countries
#| fig-align: center
#| apa-note: Estimates are the varying log-odds slopes from our multilevel model predicting perceived trustworthiness, alongside overall fixed effects (red). Points and line ranges reflect posterior medians and 66% and 95% highest posterior density intervals, respectively. Grey densities reflect samples from the posterior distribution.

tar_read(plot1_trustworthy)[[1]] +
  tar_read(plot1_trustworthy)[[2]] +
  plot_annotation(title = "How trustworthy do you think this advisor is?")
```

```{r, fig.height=3.5, fig.width=3.5}
#| label: fig-tightness
#| fig-cap: The relationship between cultural tightness and differences in perceived trustworthiness between deontological and utilitarian advisors
#| fig-align: center
#| apa-note: Positive values of the y-axis represent greater perceived trustworthiness of deontological advisors relative to utilitarian advisors. Lines and shaded areas reflect posterior medians and 50% and 95% credible intervals for visualization purposes. Points are country averages. Letters are ISO-2 country codes. The model controls for spatial and cultural non-independence between countries.

tar_read(plot8_tightness) +
  ylab(
    paste0(
      "Trustworthiness of deontological advisor/n",
      "compared to utilitarian advisor"
    )
  )
```

<br>

## Effects of Advice on Moral Judgment Updating

Given that participants generally perceive artificial moral advisors as less 
trustworthy than human advisors, we might predict that they should also be less 
likely to follow AI advice. To test this, we looked at participants' judgments 
of the moral dilemmas before and after seeing the advice, assessing whether they
updated their moral judgments in line with the advice.

```{r absolute-shifts}
# get absolute pre-post shifts
d <- 
  tar_read(absolute_diffs2) %>%
  filter(country == "Overall") %>%
  dplyr::select(!diff)

# differences in pre-post shifts between deontological and utilitarian advisors
diffs_advice <-
  d %>%
  pivot_wider(
    names_from = "advice",
    values_from = "absolute_diff"
    ) %>%
  rowwise() %>%
  mutate(diff = list(Deontological - Utilitarian)) %>%
  dplyr::select(!c(Deontological, Utilitarian))

# differences in pre-post shifts between human and AI advisors
diffs_treatment <-
  d %>%
  pivot_wider(
    names_from = "treatment",
    values_from = "absolute_diff"
    ) %>%
  rowwise() %>%
  mutate(diff = list(Human - AI)) %>%
  dplyr::select(!c(Human, AI))
```


Participants listened to the moral advice, shifting their judgments in the 
direction of the deontological or utilitarian advice that they received 
([@fig-judgment-shift]a). This was true for both artificial moral advisors and 
for human advisors, revealing a dissociation between the difference in rated 
trustworthiness and judgment updating for human vs. AI. In line with the finding
that participants perceived advisors who gave utilitarian judgments to be less 
trustworthy, the absolute size of the pre-post shifts revealed that participants
shifted their judgments more after receiving deontological advice compared to 
utilitarian advice, both when the advisor was human (difference in absolute 
shift = 
`r roundd(median(diffs_advice$diff[diffs_advice$treatment == "Human"][[1]]))`, 
95% HPDI 
[`r roundd(HPDI(diffs_advice$diff[diffs_advice$treatment == "Human"][[1]], prob = 0.95)[[1]])` 
`r roundd(HPDI(diffs_advice$diff[diffs_advice$treatment == "Human"][[1]], prob = 0.95)[[2]])`
]) and when the advisor was AI (difference in absolute shift = 
`r roundd(median(diffs_advice$diff[diffs_advice$treatment == "AI"][[1]]))`, 
95% HPDI 
[`r roundd(HPDI(diffs_advice$diff[diffs_advice$treatment == "AI"][[1]], prob = 0.95)[[1]])` 
`r roundd(HPDI(diffs_advice$diff[diffs_advice$treatment == "AI"][[1]], prob = 0.95)[[2]])`
]). By contrast, the absolute size of the pre-post shifts did not differ between
human and AI advisors, either when they gave deontological advice (difference in
absolute shift = 
`r roundd(median(diffs_treatment$diff[diffs_treatment$advice == "Deontological"][[1]]))`, 
95% HPDI 
[`r roundd(HPDI(diffs_treatment$diff[diffs_treatment$advice == "Deontological"][[1]], prob = 0.95)[[1]])` 
`r roundd(HPDI(diffs_treatment$diff[diffs_treatment$advice == "Deontological"][[1]], prob = 0.95)[[2]])`
]) or utilitarian advice (difference in absolute shift = 
`r roundd(median(diffs_treatment$diff[diffs_treatment$advice == "Utilitarian"][[1]]))`, 
95% HPDI 
[`r roundd(HPDI(diffs_treatment$diff[diffs_treatment$advice == "Utilitarian"][[1]], prob = 0.95)[[1]])` 
`r roundd(HPDI(diffs_treatment$diff[diffs_treatment$advice == "Utilitarian"][[1]], prob = 0.95)[[2]])`
]). This suggests that while people listened to advice, they particularly 
listened to deontological advice -- but it did not matter whether this came from
a human or an AI.

```{r, fig.height=5, fig.width=9.5}
#| label: fig-judgment-shift
#| fig-cap: Effects of moral advice on participant judgments
#| fig-align: center
#| apa-note: Estimated pre-post shifts in judgments on a 0-1 sliding scale, for human and AI advisors giving deontological and utilitarian advice. The grey dashed line indicates no pre-post shift in judgment. Points, line ranges, and densities reflect posterior medians, 95% highest posterior density intervals, and samples from the posterior distribution, respectively. (b) Estimated pre-post shifts in judgments, split by moral dilemma. (c) Estimated pre-post shifts in judgments, split by country. The grey dashed line indicates neutral judgments. Points and arrows reflect posterior median estimates before and after seeing the advice, respectively.

tar_read(plot_judgement_combined)
```

<br>

In additional robustness checks, the effects of advice on judgment updating were
robust across both sacrificial dilemmas, though the advice tended to be more 
persuasive in the Bike dilemma compared to the Baby dilemma 
([@fig-judgment-shift]b). We also continued to observe judgment updating when we 
restricted our analysis to judgments in the first experimental block only, 
before participants knew they would be given the opportunity to revise their 
judgments (@suppfig-judgment-shift-block1).

```{r load-int-effects2}
# get interaction effects
post <- brms::posterior_samples(tar_read(model12))
int_AI_deon <- 
  post$`bsp_motrustworthy:timepost`
int_AI_util <- 
  post$`bsp_motrustworthy:timepost` +
  post$`bsp_motrustworthy:timepost:adviceUtilitarian`
int_human_deon <- 
  post$`bsp_motrustworthy:timepost` +
  post$`bsp_motrustworthy:timepost:treatmentHuman`
int_human_util <- 
  post$`bsp_motrustworthy:timepost` +
  post$`bsp_motrustworthy:timepost:adviceUtilitarian` +
  post$`bsp_motrustworthy:timepost:treatmentHuman` +
  post$`bsp_motrustworthy:timepost:treatmentHuman:adviceUtilitarian`
```

We further explored whether judgment updating varied across countries and 
individuals. We observed similar patterns of judgment updating across all the 
countries in our sample ([@fig-judgment-shift]b). Additional moderator analyses
revealed that the size of the pre-post shift in judgments was not predicted by 
cross-cultural variables (individualism, relational mobility, cultural 
tightness) or global indices of AI progress and government readiness 
(Supplementary Figures [-@suppfig-judgment-shift-relational-mobility] - 
[-@suppfig-judgment-shift-AI-index]; @supptbl-model11). However, at the 
individual level, the amount of judgment updating did vary according to the 
perceived trustworthiness of the advisor. On average, participants updated their
judgments more when they perceived the advisor to be more trustworthy 
(@fig-judgment-shift-trustworthiness). This was true for AI advisors giving 
deontological (interaction effect = `r roundd(median(int_AI_deon))`, 95% HPDI 
[`r roundd(HPDI(int_AI_deon, prob = 0.95)[[1]])` `r roundd(HPDI(int_AI_deon, prob = 0.95)[[2]])`]) 
and utilitarian (interaction effect = `r roundd(median(int_AI_util))`, 95% HPDI 
[`r roundd(HPDI(int_AI_util, prob = 0.95)[[1]])` `r roundd(HPDI(int_AI_util, prob = 0.95)[[2]])`])
advice, and for human advisors giving deontological (interaction effect = 
`r roundd(median(int_human_deon))`, 95% HPDI
[`r roundd(HPDI(int_human_deon, prob = 0.95)[[1]])` `r roundd(HPDI(int_human_deon, prob = 0.95)[[2]])`])
and utilitarian (interaction effect = `r roundd(median(int_human_util))`, 95% HPDI
[`r roundd(HPDI(int_human_util, prob = 0.95)[[1]])` `r roundd(HPDI(int_human_util, prob = 0.95)[[2]])`])
advice. Importantly, though, even when participants gave advisors the lowest 
possible trustworthiness rating (1 on a 7-point Likert scale), they still 
continued to update their judgments in the direction of the advice that they 
received (@fig-judgment-shift-trustworthiness).

```{r, fig.height=4, fig.width=4.5}
#| label: fig-judgment-shift-trustworthiness
#| fig-cap: The relationship between perceived trustworthiness and pre-post shifts in moral judgments
#| fig-align: center
#| apa-note: Negative pre-post shifts indicate judgment updating towards the deontological option; positive pre-post shifts indicate judgment updating towards the utilitarian option. Points and line ranges reflect posterior medians and 95% highest posterior density intervals.

tar_read(plot12)
```

<br>

As well as shifting their judgments, participants' confidence in their judgment
increased after receiving the moral advice (@suppfig-confidence). This pre-post 
increase in confidence was again slightly greater after seeing deontological 
advice compared to utilitarian advice but did not differ between human and AI 
advisors. This pattern was robust across dilemmas and countries (Supplementary 
Figures [-@suppfig-confidence-baby] and [-@suppfig-confidence-bike]).

# Discussion

Some have proposed that AI tools can, and should, help us in the moral domain 
[@Giubilini2018; @Lara2020; @Sinnott2021]. But it has remained unclear whether 
and how people trust AI moral advisors, and how the kind of normative ethical 
advice that the AI provides might influence this trust. In a large-scale, 
pre-registered experiment across 12 countries, we studied people's trust in, and
deference to, artificial moral advisors who give advice aligned with different
ethical theories. We found that participants perceived AI moral advisors
as less trustworthy than humans, particularly when they gave advice based on 
utilitarian principles. Despite this relative distrust, however, people still 
listened to the moral advice from AI by updating their own moral judgments, 
especially when the advice was in line with deontological principles. 
Importantly, this judgment updating was observed regardless of whether the 
advice came from a human or an AI advisor and, perhaps concerningly, even when 
the advisor was explicitly perceived to be untrustworthy. Taken together, these
findings suggest that artificial moral advisors may play an influential role in
human morality, for better or worse.

Our first key finding is that participants perceived artificial moral advisors 
to be less trustworthy than humans. We find that even though our AI advisors 
were explicitly described as moral experts, people trusted them less than human
advisors -- even if trustworthiness was still around the midpoint of the scale.
This replicates previous work on algorithm aversion in general [@Dawes1979; 
@Dietvorst2015; @Meehl1954; @Meehl1957] and emerging findings in the moral 
domain [@Bigman2018; @Myers2025]. This pattern was observed for most of the 
countries in our sample and was not moderated by global indices of AI progress 
or government AI readiness. For those who are optimistic about the power of AMAs
to be used in assisting moral judgments [@Giubilini2018; @Sinnott2021], such 
results could lead to cautious optimism: these systems may be seen as less 
trustworthy than humans, but not completely untrustworthy. And yet, we show that
this trustworthiness is not static but depends on the specific type of moral 
judgment that the AI makes.

Our second key finding is that the type of advice drove participants' 
perceptions of the advisors. Participants trusted AI (and human) advisors who 
gave characteristically deontological advice more than advisors who gave 
utilitarian advice. These results were observed overall, found when looking at 
both dilemmas separately, and found when accounting for participants' own moral 
judgments. For AI, like for humans [@Everett2016; @Everett2018; @Everett2021], 
not all "moral" decisions equally signal trustworthiness and therefore AMAs that
are built to align with impartial utilitarian ethics may be particularly 
distrusted. Instead, artificial moral advice that aligns with deontological 
norms of integrity and respect for persons may be better received by a public 
uneasy with AI's perceived cold rationality, raising challenges to those who 
believe that utilitarianism is the correct normative standpoint by which AMAs 
should be programmed.

While the difference in perceived trustworthiness between deontological and 
utilitarian advisors generalized across most countries in our sample, we found 
that "looser" cultures (cultures with less restrictive social norms, such as 
Brazil, Mexico, and Chile) trusted deontological advisors relatively more than 
"tighter" cultures (cultures with more restrictive social norms, such as India 
and China) [@Gelfand2006]. This aligns with recent work showing that people in 
looser cultures are more averse to the use of personal force to achieve 
utilitarian ends in sacrificial dilemmas [@Couto2025]. While we did not 
pre-register any specific predictions about cross-cultural differences, there 
are several potential explanations for this association. People in tighter 
cultures tend to prioritize the prevention of negative outcomes more than looser
cultures [@Gelfand2006], which is exactly what the utilitarian decisions in our
study aimed to achieve (for example, preventing the deaths of the townspeople in
the Baby dilemma). The utilitarian decisions also involved sacrificing one 
individual for the wider group, which aligns with the focus on group cohesion in
tighter societies [@Gelfand2006]. Future research should aim to replicate this 
exploratory cross-cultural association and assess its generalizability across 
different trust measures -- notably, in our study, we found that cultural 
tightness was associated with perceived trustworthiness, but not judgment 
updating.

Our third key finding is that participants listened to the moral advice by 
updating their own moral judgments, regardless of whether the advisor was human
or AI -- but did so especially when the advice was deontological, in line with 
the effects on perceived trustworthiness. In addition, participants updated
their judgments more when the advisor was perceived to be more trustworthy, yet
they continued to update their judgments even when they explicitly perceived the
advisor as untrustworthy. This reveals a psychological paradox: people may claim
to distrust AI relative to humans, yet still update their moral views based on
AI advice. Perceptions of trustworthiness are typically understood as 
antecedents of trust [@Lee2004; @Mayer1995], even if we know that perceptions of 
AI trustworthiness do not always lead to trusting behavior [@Schlicker2025].
Yet our results highlight an underexplored and opposite pattern: that people may 
"trust" AMAs behaviorally by updating their moral judgments, while still 
perceiving them as less trustworthy than humans. This finding extends the 
growing literature on the persuasive potential of AI [@Bai2025; @Costello2024; 
@Durmus2024; @Hackenburg2024; @Hackenburg2025; @Krugel2023; @Schoenegger2025]
into the moral domain and shows that people's perceptions of AI trustworthiness
are distinct from their decisions to defer to AI in practice [@Hoff2015; 
@Lee2004; @Schlicker2025; @Vereschak2021; @Wischnewski2023]. The moral 
persuasiveness of AI is good news for those who argue for the possibility of AI 
to enhance moral judgments [@Giubilini2018; @Sinnott2021], but is likely to ring
alarm bells for those who have expressed concerns about outsourcing moral 
judgments to AI, including concerns about the lack of epistemic agency involved,
the risk of moral deskilling, and potential issues with offloading moral 
responsibility onto machines [@Landes2025; @Landes_preprint; @Liu2022; 
@Vallor2024].

For AI developers and policymakers, our judgment updating results suggest that
when testing AMA systems, developers should be cautious not to conflate
perceived trustworthiness with actual behavioral impact. At a minimum, our 
results suggest the need for safeguards for emerging AMA technologies, 
especially those operating in high-stakes domains like healthcare, military 
ethics, and criminal justice. The suggestion that people may behaviorally trust
these systems even when rating them as untrustworthy reiterates the 
importance of "designing for appropriate reliance" [@Lee2004] and ensuring 
calibrated trust [@Ferrario2024], perhaps using both trust-enhancing and 
trust-dampening cues [@Visser2025]. And yet, of course, our results here are 
descriptive and should not lead to any normative conclusions. The fact that 
people trust AMAs behaviorally by updating their moral judgments is a completely
separate question to whether people should trust AMAs, whether developers should
build AMAs at all, and whether the increasing expansion of algorithms into the 
moral domain is an appropriate vision of society [@Everett_in_prep].

Our results leave open the question of why people update their judgments in 
response to AI advice. People might be persuaded by the reasons provided by the 
AI, defer to its perceived expertise, reconsider their initial decisions when 
given the opportunity to revise them, or simply adjust their judgments to meet 
the perceived demands of the experiment. Our finding that people listened more 
to deontological than utilitarian advice suggests that our results are not 
merely due to demand characteristics. While our design was not able to tease 
apart other explanations for judgment updating, recent work has found that 
people do actively attend to the content of advice given by AMAs when updating 
their judgments, rather than blindly deferring to the advisor's perceived 
expertise [@Landes_under_review].

Our fourth key finding is that people seem to associate utilitarianism with 
being more machine-like. Utilitarianism, with its impartial focus on calculating
predicted welfare and assessing consequences, has often been criticized for the
way it seems to diminish humanity through its focus on algorithms -- that it is
a "a civilization of 'things' and not of 'persons'" [@JohnPaulII1995] and that
when utilitarians make moral decisions by rationally calculating outcomes they 
are having "one thought too many" [@Williams1981] rather than acting
instinctively based on the pull of human ties. While this characterization is 
perhaps unfair for not recognizing the impartial welfare-maximizing nature of
utilitarianism [@Kahane2018], it remains a pervasive criticism within philosophy
[@Glover1990] and in public discussions [@Torres2021]. Unsurprisingly then, with
its focus on computations of overall welfare and consequences, the utilitarian
decision-making process is sometimes seen as akin to what a machine, not a
human, would do. Indeed, there is evidence that people expect AI to be more
utilitarian [@Malle2015; @Myers2025] and here we show that people were more
surprised when deontological advice came from an AI rather than a human; that
humans who made utilitarian judgments were seen to be more machine-like than
those who made deontological judgments; and that AI systems that made
deontological judgments were seen as more human-like. This association between
utilitarianism and machine-likeness highlights the broader challenges AI systems
face in appearing not just intelligent, but *humane*. If AI developers want
artificial moral advisors to be accepted, and this is by no means a given, then
they should build systems that can do more than merely compute outcomes -- the
systems must also communicate care, character, and moral understanding in ways
that align with human moral psychology.

Future work should build on these findings in several ways. First, AMAs need to 
be aligned with a normative standard, but there are many different ethical 
frameworks that they could be aligned with. While our results cohere with 
previous work showing that utilitarian advisors are trusted less than 
deontological advisors [@Everett2018; @Everett2021; @Myers2025], future work 
should explore how AMAs are judged when their advice aligns with other 
philosophical theories, such as virtue ethics. Second, future work should assess
trust in AMAs in everyday moral situations. Sacrificial dilemmas are frequently
used in moral psychology [@Conway2013; @Duke2015; @Everett2021; @Greene2008; 
@Myers2025] due to their ability to tease apart different normative frameworks,
such as deontology and utilitarianism. This was an advantage for our study, as 
these normative frameworks are precisely the focus of discussions about how to 
ethically program AMAs [@Giubilini2018; @Landes2025; @Liu2022]. Nonetheless, 
sacrificial dilemmas have been criticized for lacking realism compared to more 
contextualized real-world scenarios [@Bauman2014]. Research should explore how 
people trust AI advisors in the kinds of everyday moral situations that LLMs are 
increasingly being used to provide advice for, such as parenting, workplace 
fairness, and relationships. These real-world contexts may trigger different 
patterns of trust and influence than abstract dilemmas. Third, future work 
should explore the role of advice consistency in the perceptions of AMAs. Since 
our design focused on individual responses from separate advisors, we were 
unable to assess the effects of advisor consistency on participants' judgments. 
However, there is reason to believe that (in)consistency could be important for
trust, especially since real AMAs are likely to be applied across many different
situations. If people want AMAs to act as dispassionate "ideal observers" 
[@Firth1952], then they may prefer AI to give consistent and unbiased moral 
judgments in line with a particular normative framework. Alternatively, people 
might prefer AI to adapt its moral decision-making dynamically depending on 
morally relevant features of the situation. Future research could employ 
multi-trial designs with different dilemmas to test these possibilities.

Overall, our findings reveal a paradox: people distrust artificial moral 
advisors relative to humans, and particularly distrust utilitarian advisors, yet
still change their judgments in response to advice from AI, even when they 
perceive the AI to be untrustworthy. This tension underscores both the promise 
and peril of using AI in the moral domain. If artificial moral advisors are to 
become more embedded in everyday life, it becomes ever more imperative that they
do more than calculate, but align with the moral intuitions, emotions, and 
values that make us human.

# Methods

## Ethical Approval

Ethical approval was granted for the study by the University of Kent's 
Psychology Research Ethics Panel (protocol ID: 202517411704229833). All 
participants provided informed consent at the beginning of the survey and were 
debriefed at the end of the survey.

## Participants and Sampling

To determine our sample size, we conducted a power simulation informed by the 
effect sizes from a previous study [@Myers2025]. This previous study found a 
small main effect of advice type (deontological vs. utilitarian) and a small
main effect of advisor type (human vs. AI) on trust. Our simulation suggested
that we would require 500 participants to detect both effects with 90% power in 
our proposed design. We aimed to recruit at least this number of participants in
each country (overall minimum N = 6000 for 12 countries) to ensure that we were
adequately powered within each country. After excluding and replacing 
participants who failed our exclusion criteria (see Exclusions), our final 
sample size was *N* = 6,896 (see @suppfig-sample for overall sample 
characteristics and @supptbl-sample-by-country for a breakdown of the sample by
country). 3,467 of these participants were randomly allocated to the AI advisor
condition and 3,429 participants were randomly allocated to the human advisor
condition. These data were collected over a 23-day period from 8 April to 1 May
2025.

We recruited participants from 12 countries -- Brazil, Chile, China, France, 
Germany, India, Mexico, Poland, South Africa, Turkey, the United Kingdom, and 
the United States -- through the online panel aggregator Prime Panels 
(<https://www.cloudresearch.com/products/prime-panels/>; 
[@fig-study-overview]c). We chose this set of countries to maximize variation in
geographic spread, cultural backgrounds, and levels of AI progress according to
global indices (@suppfig-cross-cultural). We used quotas to target 
nationally representative samples with respect to age and gender, relaxing these
quotas when data collection slowed in the final 15% of participants in each 
country. Despite relaxing the quotas, we found that the differences between 
observed and targeted proportions in each quota category were less than 10% for
all countries except China, which was skewed slightly towards younger 
participants (@suppfig-sample-representativeness).

## Design

The general flow of the experiment is presented in [@fig-study-overview]a. The 
experiment followed a mixed design. We randomized participants to see advice 
from either human advisors or AI advisors throughout the whole experiment 
(between-subjects). We then presented participants with two experimental blocks
in a within-subjects design. In the first block, participants saw either
deontological or utilitarian advice (randomly counterbalanced) for either the
Baby dilemma or the Bike dilemma (randomly counterbalanced; see Procedure for
more details about the dilemmas). In the second block, participants then saw the
other direction of advice for the other moral dilemma.

## Procedure

Participants completed the survey experiment in Qualtrics 
(<https://www.qualtrics.com/>). At the beginning of the survey, participants 
chose their survey language, completed a Captcha verification question, provided
informed consent, reported whether they currently lived in the country of 
intended recruitment, provided their age and gender, and answered two attention
check questions. Participants were unable to proceed with the survey if they met
any of the following criteria: (1) they had a Captcha score of less than 0.5,
(2) they did not provide informed consent, (3) they stated that they lived in a
different country to that of intended recruitment, (4) they were in an
age/gender quota group that had already been filled, or (5) they failed either
of our two attention check questions.

If participants proceeded with the survey, we presented them with the following
study preamble text: "In life, there can be difficult situations that require 
people to make moral decisions. To make these decisions, we often turn to 
external experts for help." In the human condition, the text continued: "Moral 
Advisors draw on their extensive ethical and philosophical training to provide 
recommendations in difficult moral cases." In the AI condition, the text 
continued: "Artificial Moral Advisors draw on the latest advancements in 
artificial intelligence to provide recommendations in difficult moral cases. 
These systems employ advanced statistical and machine learning techniques to 
engage in human-level moral reasoning and ethical problem solving."

We then presented participants with the two experimental blocks. At the start of
each block, we presented participants with the sacrificial dilemma (see 
Supplementary Methods for full dilemma wordings). The Baby dilemma described a
situation where townspeople are hiding from enemy soldiers, and someone must
choose between smothering their crying baby to save everyone or letting the baby
cry and risking everyone's death. The Bike dilemma described a situation where a
motorcycle rider is about to crash on the road, and someone must choose between
causing the rider's likely death by forcing them off the road or allowing a
crash that could result in the deaths of many other riders behind them. These
dilemmas were replicated from Myers and Everett's [@Myers2025] Study 2. On the 
same page as the dilemma, we asked participants for their own judgment on what
they thought should be done in the dilemma, rated on a 0-1 sliding scale from 
deontological judgment (not smothering the baby, not crashing the rider) to 
utilitarian judgment (smothering the baby, crashing the rider). We also asked 
participants how confident they were in their judgment on a 1-7 Likert scale. 
Participants' responses to these measures suggested that they interpreted the 
0-1 sliding scale as the relative direction and certainty of their moral 
convictions, with distance from the midpoint of this scale correlating 
positively with confidence ratings (*r* = 
`r roundd(cor(abs(tar_read(data)$judgement_pre - 0.5), tar_read(data)$confidence_pre, use = "pairwise"))`).

On the following page, we presented participants with advice about the dilemma 
from a moral advisor. In the human condition, we told participants that 
Dr. Johnson or Dr. Smith (names randomly counterbalanced across blocks) drew on 
their extensive ethical and philosophical training. In the AI condition, we told 
participants that ETHIC-AI or VIRTUE-BOT (names counterbalanced) drew on 
advancements in machine learning about moral cases. We then presented the moral
advice ([@fig-study-overview]b). When giving deontological advice, the advisor
stated that "an important principle in ethics is that killing people is just 
wrong, and this duty to not kill should apply even if killing has good 
consequences in a specific case." When giving utilitarian advice, the advisor 
stated that "an important principle in ethics is to think about the greater 
good, and in this specific case killing the one person would bring about better
consequences overall." On the same page as the advice, we then asked 
participants the following questions about their perceptions of the advisor on 
1-7 Likert scales:

- How trustworthy do you think [the advisor] is?
- How much would you blame someone if they followed [the advisor's] 
recommendation?
- Based on their advice, how willing would you be to trust [the advisor] on 
other issues?
- How surprised were you at [the advisor's] recommendation?
- How well do the following words describe [the advisor]? Machine-like ↔ 
human-like

<br>

On the next page, we reminded participants about the moral advice and then asked
them again to judge what they thought should be done (0-1 sliding scale) and 
report how confident they were in their judgment (1-7 Likert). Then on the final
page of the block, we asked participants a binary-choice comprehension question
("What did the advisor recommend in this dilemma?") to check that they could 
correctly recall the deontological or utilitarian advice. After seeing the first 
experimental block, participants repeated this process in a second experimental
block with the other advisor giving the other direction of advice for the other
moral dilemma.

After the two experimental blocks, we asked participants some general questions
about "AI tools like ChatGPT", including how familiar they were with AI tools,
how frequently they used AI tools, and how trustworthy they thought AI tools
were. We then asked participants several demographic questions, including their
highest education level, their subjective social status (measured with the
MacArthur ladder), their political ideology (from left-wing to right-wing; not
asked in China), and their religiosity. We then debriefed participants about the
purpose of the study and gave them an open-ended form to leave any comments or
remarks about the study.

## Translations

In non-English-speaking countries, we translated the survey using a forward- and 
back-translation procedure. First, a native speaker translated the survey from 
English into the target language. Then, another native speaker (who had not yet
seen the survey) translated the survey back into English. Any discrepancies were
discussed between forward- and back-translators and any issues were resolved.
All translators then checked the final Qualtrics surveys to ensure that the
translations had been implemented correctly.

## Pre-registration

We pre-registered our hypotheses, study design, sampling plan, exclusion
criteria, and analysis plan on the Open Science Framework on 8 April 2025 
(<https://osf.io/rcufb>). We deviated from the pre-registration by using 95% 
HPDIs as the inference criteria rather than 95% equal-tailed credible intervals,
as suggested by a reviewer. Since the results were identical when using either
type of interval, we report HPDIs only in the paper.

## Exclusions

We excluded and replaced participants who completed the survey but met any of 
the following criteria: (1) they had taken the survey more than once as 
indicated by duplicate IP addresses or assignment IDs, (2) they straight-lined
on both of the advisor-perceptions survey pages by providing the same responses
to all items, (3) they sped through the survey, either by completing the survey
in less than 100 seconds (~2 seconds per question/text) or completing the survey
at least two median absolute deviations faster than the median response time,
(4) they took more than 24 hours to complete the survey, (5) they failed to
answer more than 50% of the questions in the survey, or (6) they wrote
nonsensical responses in our open-ended question, as independently judged by
translators who were blind to experimental condition and responses to our
primary measures.

We also excluded responses to moral dilemmas if the participant failed the 
comprehension question for that moral dilemma. This exclusion strategy resulted
in our final sample size of 6,896 participants. However, in additional 
"intention-to-treat" analyses, we retained these comprehension failures,
resulting in a larger sample size of 7,096 participants.

## Statistical Analysis

To analyse the perceptions of the advisors, we fitted Bayesian cumulative-link
ordinal models to the Likert scale data, modelling the interaction between
advice type (deontological vs. utilitarian) and advisor type (human vs. AI). To
split by moral dilemma, we included dilemma (Baby vs. Bike) as an additional
moderator. In exploratory models, we interacted one or both of the experimental
effects with individual-level characteristics (religiosity, political ideology, 
familiarity with AI, frequency of AI use) or cultural-level variables
(relational mobility [@Thomson2018], tightness-looseness [@Eriksson2021], 
individualism [@Hofstede2010], the global AI index [@TortoiseMedia], 
AI readiness [@OxfordInsights]). These models were fitted separately as we did 
not have a clear causal model guiding the inclusion of controls or cross-level 
interactions within a single model [@Cinelli2024; @McElreath2020].

To analyse participants' judgments and confidence, we fitted Bayesian 
zero-one-inflated beta models to the 0-1 slider data and Bayesian 
cumulative-link ordinal models to the Likert scale data, modelling the
interaction between time (pre- vs. post- advice), advice type (deontological vs. 
utilitarian), and advisor type (human vs. AI). To split by moral dilemma, we 
included dilemma (Baby vs. Bike) as an additional moderator. In exploratory 
models, we interacted the effect of time with cultural-level variables 
(relational mobility, tightness-looseness, individualism, the global AI index, 
AI readiness) and participants' trustworthiness ratings for the advisor. These 
models were fitted separately.

For all models, we included varying intercepts for participants and varying 
intercepts and slopes for countries (see Supplementary Methods for full model
specifications). When analyzing the moderating effects of cultural variables, we 
additionally controlled for the non-independence of countries by allowing the 
country-level varying effects to covary according to geographic and linguistic 
proximity matrices constructed in previous research [@Claessens2023].

We used regularizing priors centered on zero for all models to impose 
conservatism on parameter estimates, constraining the models to be skeptical of 
unrealistically large relationships while allowing the data to inform the 
direction of the posterior. These priors were chosen using prior predictive 
checks (@suppfig-prior-check). However, our results held when we used a wider
prior on the experimental effects (@suppfig-means-prior-sensitivity; 
@supptbl-contrasts-prior-sensitivity).

All models converged normally ($\hat{R}$ ≤ 1.01). See Supplementary Figures 
[-@suppfig-trace-trustworthy] - [-@suppfig-trace-judgements] for MCMC trace 
plots and Supplementary Figures [-@suppfig-pp-check-ordinal] and 
[-@suppfig-pp-check-judgements] for posterior predictive checks. We used 95% 
highest posterior density intervals as the inference criteria.

## Reproducibility

We conducted all analyses in R version 4.4.2 [@RCoreTeam]. We fitted Bayesian
regression models using the *brms* [@Burkner2017] package. We produced 
visualizations using the *ggdist* [@Kay2024], *ggplot2* [@Wickham2016], and 
*patchwork* [@Pedersen2025] packages. We reproducibly generated the manuscript 
using the *targets* [@Landau2021] package and *quarto* [@Allaire2024]. All data,
materials, and code to reproduce the analyses and figures in this paper can be 
found here: <https://github.com/ScottClaessens/trustAMAs>

\newpage

# Acknowledgements

This work was generously supported by funding from the Economic and Social 
Research Council (ES/V015176/1), a Philip Leverhulme Prize (PLP-2021-095), and
the Horizon Europe UK Guarantee via the UKRI (EP/Y00440X/1) awarded to JACE.

# Data and Code Availability

All data and original code can be found here:
<https://github.com/ScottClaessens/trustAMAs>

# Statement of Interests

The authors have no conflicts of interest to disclose.

\newpage

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

\newpage

{{< include appendix.qmd >}}
